---
title: "Case Study - SportsAnalytics"
author: "Shivam Batra"
date: "31/10/2021"
output:
        pdf_document: default
        html_document: default
---
## ----setup------------------------------------------------
```{r}
knitr::opts_chunk$set(tidy = FALSE, highlight = TRUE, comment = NA, 
                      prompt = FALSE, fig.width = 6, fig.height = 3.5,
                      message = FALSE)
options(digits = 3)
options(width = 64)
options(continue = " ")

library(tidyverse)
library(dplyr)
```


## ------------------------------------------------------------------------
# Dataset Used
# Baseball  data for the past 10 years
```{r}
spahn <- read.csv("C:\\Users\\hp\\Desktop\\datasets\\spahn.csv")
spahn %>% slice(1:3) 
```
## ------------------------------------------------------------------------
```{r}
spahn %>% slice(1:10) %>% select(Age, W, L, ERA)
```
## ------------------------------------------------------------------------
```{r}
spahn %>%
  summarize(LO = min(ERA), 
            QL = quantile(ERA, .25), QU = quantile(ERA, .75),
            M = median(ERA), HI = max(ERA))
```
## ------------------------------------------------------------------------
```{r}
spahn %>% filter(ERA == min(ERA)) %>% select(Age)
```
## ------------------------------------------------------------------------
```{r}
spahn %>% 
  mutate(FIP = (13 * HR + 3 * BB - 2 * SO) / IP) -> spahn
```
## ------------------------------------------------------------------------
```{r}
spahn %>% 
  arrange(FIP) %>% 
  select(Year, Age, W, L, ERA, FIP) %>%
  head()
```
## ------------------------------------------------------------------------
```{r}
spahn %>% filter(Tm == "BSN" | Tm == "MLN") -> spahn1
```
## ------------------------------------------------------------------------
```{r}
spahn1 <- spahn1 %>%
  mutate(Tm = factor(Tm, levels = c("BSN", "MLN")))
```
## ------------------------------------------------------------------------
```{r}
spahn1 %>% 
  group_by(Tm) %>% 
  summarize(mean_W.L = mean(W.L, na.rm = TRUE),
            mean_ERA = mean(ERA),
            mean_WHIP = mean(WHIP),
            mean_FIP = mean(FIP))
```
## ------------------------------------------------------------------------
```{r}
NLbatting <- read_csv("C:\\Users\\hp\\Desktop\\datasets\\NLbatting.csv")
ALbatting <- read_csv("C:\\Users\\hp\\Desktop\\datasets\\ALbatting.csv")
batting <- bind_rows(NLbatting, ALbatting)
```
## ------------------------------------------------------------------------
```{r}
NLpitching <- read_csv("C:\\Users\\hp\\Desktop\\datasets\\NLpitching.csv")
NL <- inner_join(NLbatting, NLpitching, by = "Tm")
```
## ------------------------------------------------------------------------
```{r}
NLbatting %>% filter(HR > 150) -> NL_150
```
## ------------------------------------------------------------------------
```{r}
W <- c(8, 21, 15, 21, 21, 22, 14)
L <- c(5, 10, 12, 14, 17, 14, 19)
L
```
## ------------------------------------------------------------------------
```{r}
Win.Pct <- 100 * W / (W + L)
Win.Pct
```
## ------------------------------------------------------------------------
```{r}
Year <- seq(from = 1946, to = 1952)
Year
```
## ------------------------------------------------------------------------
```{r}
Year <- 1946 : 1952
Age <- Year - 1921
```
## ----spahngraph, fig.cap="Scatterplot of the winning percentage against age for Warren Spahn's seasons playing for the Boston Braves."----
```{r}
plot(Age, Win.Pct)
```
## ------------------------------------------------------------------------
```{r}
mean(Win.Pct)
```
## ------------------------------------------------------------------------
```{r}
100 * sum(W) / (sum(W) + sum(L))
```
## ------------------------------------------------------------------------
```{r}
sort(W)
```
## ------------------------------------------------------------------------
```{r}
cumsum(W)
```
## ------------------------------------------------------------------------
```{r}
summary(Win.Pct)
```
## ------------------------------------------------------------------------
```{r}
W[c(1, 2, 5)]
```
## ------------------------------------------------------------------------
```{r}
W[1 : 4]
```
## ------------------------------------------------------------------------
```{r}
W[-c(1, 6)]
```
## ------------------------------------------------------------------------
```{r}
Win.Pct > 60
```
## ------------------------------------------------------------------------
```{r}
(W > 20) & (Win.Pct > 60)
```
## ------------------------------------------------------------------------
```{r}
Win.Pct == max(Win.Pct)
```
## ------------------------------------------------------------------------
```{r}
Year[Win.Pct == max(Win.Pct)]
```
## ------------------------------------------------------------------------
```{r}
Year[W + L > 30]
```
## ------------------------------------------------------------------------
```{r}
Year <- 2008 : 2017
NL <- c("PHI", "PHI", "SFN", "SLN", "SFN",
        "SLN", "SFN", "NYN", "CHN", "LAN")
AL <- c("TBA", "NYA", "TEX", "TEX", "DET",
        "BOS", "KCA", "KCA", "CLE", "HOU")
Winner <- c("NL", "AL", "NL", "NL", "NL",
            "AL", "NL", "AL", "NL", "AL")
N_Games <- c(5, 6, 5, 7, 4, 7, 7, 5, 7, 7)
```
## ------------------------------------------------------------------------
```{r}
WS_results <- tibble(
  Year = Year, NL_Team = NL, AL_Team = AL,
  N_Games = N_Games, Winner = Winner)
WS_results
```
## ------------------------------------------------------------------------
```{r}
grep("NY", c(AL, NL), value = TRUE)
```
## ------------------------------------------------------------------------
```{r}
WS_results %>% 
  group_by(Winner) %>% 
  summarize(N = n()) -> WS
WS
```
## ----barplot, fig.cap="Bar graph of the number of wins of the American League and National League teams in the World Series between 2003 and 2012."----
```{r}
ggplot(WS, aes(x = Winner, y = N)) + 
  geom_col()
```
## ----include=FALSE-------------------------------------------------------
```{r}
ggplot(WS_results, aes(x = Winner)) + 
  geom_bar()
```
## ------------------------------------------------------------------------
```{r}
WS_results %>% 
  group_by(NL_Team) %>% 
  summarize(N = n())
```
## ------------------------------------------------------------------------
```{r}
WS_results <- WS_results %>%
  mutate(NL_Team = factor(NL_Team, 
                          levels = c("NYN", "PHI", "CHN", 
                                     "SLN", "LAN", "SFN")))
```
## ------------------------------------------------------------------------
```{r}
str(WS_results$NL_Team)
```
## ------------------------------------------------------------------------
```{r}
WS_results %>% 
  group_by(NL_Team) %>% 
  summarize(N = n())
```
## ------------------------------------------------------------------------
```{r}
world_series <- list(Winner = Winner, Number.Games = N_Games,
                     Seasons = "2008 to 2017")
```
## ------------------------------------------------------------------------
```{r}
world_series$Number.Games
```
## ------------------------------------------------------------------------
```{r}
world_series[[2]]
```
## ------------------------------------------------------------------------
```{r}
pluck(world_series, "Number.Games")
```
## ------------------------------------------------------------------------
```{r}
world_series["Number.Games"]
```
## ------------------------------------------------------------------------
```{r}
WS_results$NL_Team
pull(WS_results, NL_Team)
```
## ------------------------------------------------------------------------
```{r}
hr_rates <- function(age, hr, ab) {
  rates <- round(100 * hr / ab, 1)
  list(x = age, y = rates)
}
```
## ------------------------------------------------------------------------
```{r}
HR <- c(13, 23, 21, 27, 37, 52, 34, 42, 31, 40, 54)
AB <- c(341, 549, 461, 543, 517, 533, 474, 519, 541, 527, 514)
Age <- 19 : 29
hr_rates(Age, HR, AB)
```
## ------------------------------------------------------------------------
```{r}
HR <- c(13, 23, 21, 27, 37, 52, 34, 42, 31, 40, 54)
AB <- c(341, 549, 461, 543, 517, 533, 474, 519, 541, 527, 514)
Age <- 19 : 29
hr_rates <- hr_rates(Age, HR, AB)
Mantle <- data.frame(Age, HR, AB, Rates = hr_rates$y)
```
## ------------------------------------------------------------------------
```{r}
write_csv(Mantle, "C:\\Users\\hp\\Desktop\\datasets\\mantle.csv")
```
## ------------------------------------------------------------------------
```{r}
list.files("data", pattern = "mantle")
```
## ------------------------------------------------------------------------
```{r}
library(Lahman)
```
## ------------------------------------------------------------------------
```{r}
Batting %>% 
  filter(yearID >= 1960, yearID <= 1969) ->  Batting_60
```
## ------------------------------------------------------------------------
```{r}
Batting_60 %>% 
  group_by(playerID) %>% 
  summarize(HR = sum(HR)) -> hr_60
```
## ------------------------------------------------------------------------
```{r}
hr_60 %>% 
  arrange(desc(HR)) -> hr_60
head(hr_60)
```
## ------------------------------------------------------------------------
```{r}
Batting %>% 
  filter(yearID >= 1960, yearID <= 1969) %>%
  group_by(playerID) %>% 
  summarize(HR = sum(HR)) %>%
  arrange(desc(HR)) %>%
  head()
```
## ------------------------------------------------------------------------
```{r}
hr_leader <- function(data) {
  data %>% 
    group_by(playerID) %>% 
    summarize(HR = sum(HR)) %>%
    arrange(desc(HR)) %>%
    head(1)
}
```
## ------------------------------------------------------------------------
```{r}
Batting %>%
  mutate(decade = 10 * floor(yearID / 10)) %>%
  split(pull(., decade)) %>%
  map_df(hr_leader, .id = "decade")
```
## ------------------------------------------------------------------------
```{r}
Batting %>% 
  group_by(playerID) %>% 
  summarize(tAB = sum(AB, na.rm = TRUE),
            tHR = sum(HR, na.rm = TRUE),
            tSO = sum(SO, na.rm = TRUE)) -> long_careers
```
## ------------------------------------------------------------------------
```{r}
Batting_5000 <- filter(long_careers, tAB >= 5000)
## ------------------------------------------------------------------------
head(Batting_5000)
```
## ----hrso, fig.cap="Scatterplot of the homerun rates and strikeout rates of all players with at least 5000 career at-bats. A smoothing curve is added to the plot to show that home run rates and strikeout rates have a positive association."----
```{r}
#  Q. plot the SO/AB (Y) against the HR/AB (X)
ggplot(Batting_5000, aes(x = tHR / tAB, y = tSO / tAB)) +
   geom_point() + geom_smooth()
```
## ------------------------------------------------------------------------
```{r}
career.pitching <- Pitching %>% 
  group_by(playerID) %>% 
  summarize(SO = sum(SO, na.rm = TRUE),
            BB = sum(BB, na.rm = TRUE),
            IPouts = sum(IPouts, na.rm = TRUE),
            midYear = median(yearID, na.rm = TRUE)) 
```
## ------------------------------------------------------------------------
#Question 1
#In R, place the stolen base, caught stealing, and game counts in the vectors SB, CS, and G.
# Import the Hall of Fame data set from the repo:
```{r}
hof <- read_csv("https://raw.githubusercontent.com/maxtoki/baseball_R/master/data/hofbatting.csv", col_names = TRUE)
dim(hof)
```
## ------------------------------------------------------------------------
```{r}
head(hof)
```
## ------------------------------------------------------------------------
```{r}
SB <- hof %>% select(SB) %>% arrange(desc(SB))
CS <- hof %>% select(CS) %>% arrange(desc(CS))
G <- hof %>% select(G) %>% arrange(desc(G))
```


```{r}

# For all players, compute the number of stolen base attempts SB + CS and store in the vector SB.Attempt.
SB.Attempt <- hof %>% select(SB, CS) %>% transmute(SB.Attempt = SB + CS) %>% arrange(desc(SB.Attempt))
head(SB.Attempt)
```
## ------------------------------------------------------------------------
# For all players, compute the success rate Success.Rate = SB / SB.Attempt.
```{r}
Success.Rate <- hof %>% select(SB, CS) %>% transmute(Success.Rate = SB / (SB + CS)) %>% arrange(desc(Success.Rate))
head(Success.Rate)
```
## ------------------------------------------------------------------------
# Compute the number of stolen bases per game SB.Game = SB / Game.
```{r}
SB.Game <- hof %>% select(SB, G) %>% transmute(SB.Game = SB / G) %>% arrange(desc(SB.Game))
head(SB.Game)
```
## --------------------------------------------------------------------------
## --------------------------------------------------------------------------
# 2. Question 2 - Suppose one records the outcomes of a batter in ten plate appearances: *Single*, *Out*, *Out*, *Single*, *Out*, *Double*, *Out*, *Walk*, *Out*, *Single* + Use the c function to collect these outcomes in a character vector "outcomes."
```{r}
outcomes <- c("Single", "Out", "Out", "Single", "Out", "Double", "Out", "Walk", "Out", "Single")
table(outcomes)
```
## --------------------------------------------------------------------------
# Question 3
# In R, place the wins and losses in the vectors W and L, respectively. Also, create a character vector Name containing the last names of these pitchers.
```{r}
Pitching <- read_csv("https://raw.githubusercontent.com/maxtoki/baseball_R/master/data/pitching.csv", col_names = TRUE)
dim(Pitching)
```
## --------------------------------------------------------------------------
```{r}
head(Pitching)
```
## --------------------------------------------------------------------------
```{r}
pitching.350 <- Pitching %>% 
  group_by(playerID) %>% 
  summarize(W = sum(W), L = sum(L), SO = sum(SO), BB = sum(BB)) %>% 
  filter(W >= 350) %>% 
  rename(Name = playerID)
dim(pitching.350)
```
## --------------------------------------------------------------------------
```{r}
head(pitching.350)
```
## --------------------------------------------------------------------------
# Compute the winning percentage for all pitchers defined by 100 × W/(W+L) and put these winning percentages in the vector Win.PCT.
```{r}
Win.PCT <- pitching.350 %>% 
  mutate(Win.PCT = ((100 * W) / (W + L))) %>% 
  select(Name, Win.PCT)
Win.PCT
```
## ---------------------------------------------------------------------------
# By use of the command Wins.350 <- data.frame(Name, W, L, Win.PCT) create a data frame Wins.350 containing the names, wins, losses, and winning percentages.
```{r}
Wins.350 <- pitching.350 %>% 
  select(Name, W, L) %>% 
  inner_join(Win.PCT, by = "Name")
Wins.350
```
## ---------------------------------------------------------------------------
# By use of the "order" function, sort the data frame Wins.350 by winning percentage. Among these pitchers, who had the largest and smallest winning percentages?
```{r}
Wins.350 %>% arrange(desc(Win.PCT))
```
## ---------------------------------------------------------------------------
# Question 4
# Compute the strikeout-walk ratio by SO/BB and put these ratios in the vector SO.BB.Ratio.
```{r}
SO.BB.Ratio <- pitching.350 %>% 
  mutate(SO.BB.Ratio = (SO / BB)) %>% 
  select(Name, SO.BB.Ratio)
SO.BB.Ratio
```
## ---------------------------------------------------------------------------
# by use of the command "SO.BB <- data.frame(Name, SO, BB, SO.BB.Ratio)" create a data frame "SO.BB" containing the names, strikeouts, walks, and strikeout-walk ratios.
```{r}
SO.BB <- pitching.350 %>% 
  mutate(SO.BB.Ratio = (SO / BB)) %>% 
  select(Name, SO, BB, SO.BB.Ratio)
SO.BB
```
## ---------------------------------------------------------------------------
# By use of the subset function, find the pitchers who had a strikeout-walk ratio exceeding 2.8.
```{r}
SO.BB %>% filter(SO.BB.Ratio > 2.8)
```
## ---------------------------------------------------------------------------
# By use of the order function, sort the data frame by the number of walks. Did the pitcher with the largest number of walks have a high or low strikeout-walk ratio?
```{r}
SO.BB %>% arrange(desc(BB))
```
## ---------------------------------------------------------------------------
# Question 5
# computes the cumulative strikeouts, cumulative walks, mid career year, and the total innings pitched (measured in terms of outs) for a pitcher whose season statistics are stored in the data frame d.
```{r}
stats <- function(d){
  c.SO <- sum(d$SO, na.rm=TRUE)
  c.BB <- sum(d$BB, na.rm=TRUE)
  c.IPouts <- sum(d$IPouts, na.rm=TRUE)
  c.midYear <- median(d$yearID, na.rm=TRUE)
  data.frame(SO = c.SO, BB = c.BB, IPouts = c.IPouts,
  midYear=c.midYear)
}
```
## ---------------------------------------------------------------------------
```{r}
career.pitching <- Pitching %>% group_by(playerID) %>% summarize(SO = sum(SO, na.rm = TRUE), BB = sum(BB, na.rm = TRUE), IPouts = sum(IPouts, na.rm = TRUE), midyear = median(yearID, na.rm = TRUE))
head(career.pitching)
```
## ----------------------------------------------------------------------------
#References
```{r}
# https://www.crcpress.com/Analyzing-Baseball-Data-with-R-Second-Edition/Marchi-Albert-Baumer/p/book/9780815353515
```

